### 2016-11-15 write-ups/PLDI/paper.tex
Finished pass over implementation.

%
% LaTeX template for prepartion of submissions to PLDI'16
%
% Requires temporary version of sigplanconf style file provided on
% PLDI'16 web site.
% 
\documentclass[numbers,nocopyrightspace,10pt]{sigplanconf}
% \documentclass[pldi-cameraready]{sigplanconf-pldi16}

%
% the following standard packages may be helpful, but are not required
%
\usepackage{SIunits}            % typset units correctly
\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{url}                  % format URLs
\usepackage{listings}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
% known bug: http://tex.stackexchange.com/questions/1522/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink

\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{calc}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[T1]{fontenc}
\usepackage{listings}
\usepackage{color}
\usepackage{xspace}
\usepackage{array}
\usepackage{multirow}
\usepackage{varwidth}
\usepackage{tabularx}  % for 'tabularx' environment and 'X' column type
\usepackage{ragged2e}  % for '\RaggedRight' macro (allows hyphenation)
\usepackage{xcolor,colortbl}
\usepackage{graphicx}
\usepackage{mathpartir}

\usepackage{booktabs}
\usepackage{subfig}

\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}   % print a hyperlinked DOI

\newcommand\crule[3][black]{\textcolor{#1}{\rule{#2}{#3}}}

%\newcommand{\todo}[1]{}
%\renewcommand{\todo}[1]{{\color{red} TODO: {#1}}}
\newcolumntype{C}[1]{>{\centering}m{#1}}
%\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\newcommand{\ADComment}[1]{\textcolor{green}{AD: #1}}
%\newcommand{\TSComment}[1]{\textcolor{blue}{TS: #1}}
\newcommand{\HEComment}[1]{\textcolor{orange}{HE: #1}}
%\newcommand{\COULDCUT}[1]{\textcolor{blue}{#1}}
\newcommand{\TSComment}[1]{\textcolor{purple}{TS: #1}}

\newcommand{\KC}{\mathit{KC}}

\newcommand{\kernelmerge}{\textsc{KernelMerge}}

\newcommand{\myfig}{Fig.~}
\newcommand{\myfiglong}{Figure~}
\newcommand{\myfigs}{Figs.~}
\newcommand{\mytab}{Tab.~}
\newcommand{\mytablong}{Table~}
\newcommand{\mysec}{Sec.~}
\newcommand{\myalg}{Alg.~}
\newcommand{\nvidia}{Nvidia\xspace}
\newcommand{\totalcombinations}{70\xspace}
\newcommand{\positivecombinations}{55\xspace}
\newcommand{\cbedot}{\emph{cbe-dot}\xspace}
\newcommand{\cbeht}{\emph{cbe-ht}\xspace}
\newcommand{\ctoctree}{\emph{ct-octree}\xspace}
\newcommand{\tpotm}{\emph{tpo-tm}\xspace}
\newcommand{\sdkred}{\emph{sdk-red}\xspace}
\newcommand{\sdkrednf}{\emph{sdk-red-nf}\xspace}
\newcommand{\lsbh}{\emph{ls-bh}\xspace}
\newcommand{\lsbhnf}{\emph{ls-bh-nf}\xspace}
\newcommand{\cubscan}{\emph{cub-scan}\xspace}
\newcommand{\cubscannf}{\emph{cub-scan-nf}\xspace}
\newcommand{\heuristiccombinations}{eight\xspace}
\newcommand{\effectivethreshold}{5\%\xspace}
\definecolor{Gray3}{gray}{0.9}
\definecolor{Gray2}{gray}{0.75}
\definecolor{Gray1}{gray}{0.6}

% lstlisting macros
\lstset{
  language=C, % choose the language of the code
  showspaces=false, % show spaces adding particular underscores
  showstringspaces=false, % underline spaces within strings
  showtabs=false, % show tabs within strings adding particular underscores
  tabsize=8, % sets default tabsize to 2 spaces
  captionpos=b, % sets the caption-position to bottom
  mathescape=true, % activates special behaviour of the dollar sign
  basicstyle=\scriptsize\tt,
  columns=fullfelxible,
  %commentstyle=\rmfamily\itshape,
  morekeywords={barrier,kernel,global,__device__,__syncthreads,__global__,bool,then,transmit},
  escapeinside={(*@}{@*)},
  numbers=left
}

\newcommand{\transmit}{\mathsf{transmit}}

\newcommand{\code}[1]{\lstset{basicstyle=\tt}\lstinline!#1!\lstset{basicstyle=\scriptsize\tt}}

\newcommand{\MP}{\textsf{MP}}
\newcommand{\LB}{\textsf{LB}}
\newcommand{\SB}{\textsf{SB}}
\newcommand{\noisethreshold}{3\xspace}
\newcommand{\TotalLitmusTests}{196.6\xspace}
\newcommand{\loadop}{\textsf{ld}\xspace}
\newcommand{\storeop}{\textsf{st}\xspace}

\newcommand{\Td}[2]{{#1}_{#2}}
\newcommand{\Tdl}[3]{\langle \Td{#1}{#2}, #3 \rangle}
\newcommand{\Tdlsigma}[4]{\Tdl{#1}{#2}{#4@#3}}

\newcommand{\randstrsolo}{\emph{rand-str}\xspace}
\newcommand{\sysstrsolo}{\emph{sys-str}\xspace}
\newcommand{\nostrsolo}{\emph{no-str}\xspace}
\newcommand{\cachestrsolo}{\emph{cache-str}\xspace}

\newcommand{\randstr}[1]{\emph{rand-str}#1}
\newcommand{\sysstr}[1]{\emph{sys-str}#1}
\newcommand{\nostr}[1]{\emph{no-str}#1}
\newcommand{\cachestr}[1]{\emph{cache-str}#1}

\newcommand{\NumGPUs}{seven\xspace}
\newcommand{\NumApplications}{ten\xspace}
\newcommand{\NumApplicationsUnique}{seven\xspace}

\newcommand{\offerfork}{\mathsf{offer\_fork}}
\newcommand{\offerkill}{\mathsf{offer\_kill}}
\newcommand{\globalbarrier}{\mathsf{global\_barrier}}
\newcommand{\resizingglobalbarrier}{\mathsf{resizing\_global\_barrier}}
\newcommand{\getgroupid}{\mathsf{get\_group\_id}}
\newcommand{\getnumgroups}{\mathsf{get\_num\_groups}}
\newcommand{\getlocalid}{\mathsf{get\_local\_id}}
\newcommand{\getglobalid}{\mathsf{get\_global\_id}}
\newcommand{\getlocalsize}{\mathsf{get\_local\_size}}
\newcommand{\getglobalsize}{\mathsf{get\_global\_size}}

\newcommand{\keyword}[1]{\mathsf{#1}}

\newcommand{\NumAlgorithms}{8}

\begin{document}

\title{Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs}

%
% any author declaration will be ignored  when using 'pldi' option (for double blind review)
%

\authorinfo{}
{\makebox{} \\
}
{}

%\authorinfo{Tyler Sorensen}
%{\makebox{Imperial College London, UK} \\
%}
%{t.sorensen15@imperial.ac.uk}

%\authorinfo{Alastair F. Donaldson}
%{\makebox{Imperial College London, UK} \\
%}
%{alastair.donaldson@imperial.ac.uk}

\maketitle

\vspace{-3mm}
\begin{abstract}
There is growing interest in accelerating irregular data-parallel
algorithms on GPUs.  These algorithms are typically \emph{blocking},
so require fair scheduling.  But GPU programming models (e.g.\ OpenCL)
do not mandate fair scheduling, and GPU schedulers are inherently
unfair in practice.  Current approaches avoid this issue by exploiting
scheduling quirks particular to today's GPUs in a manner that does not
allow the GPU to be shared with other workloads (such as graphics
rendering tasks).  We propose \emph{cooperative kernels}, an extension
to the traditional GPU programming model geared towards writing
blocking algorithms.  Workgroups of a cooperative kernel \emph{are}
fairly scheduled, and multitasking is supported via a small set of
language extensions that the kernel and scheduler use to cooperate.
We describe the semantics of the cooperative kernels programming model
and our prototype implementation.  We evaluate the approach by porting
a set of irregular work stealing and graph algorithms to our
programming model and examining their performance under multitasking
across three GPUs.  Our implementation exploits no
vendor-specific hardware, driver or compiler support, so our encouraging
results provide a lower-bound on the efficiency with
which cooperative kernels can be implemented in practice.

\end{abstract}

\section{Introduction}\label{sec:intro}

\paragraph{The needs of irregular data-parallel algorithms}
%Acceleration of general-purpose computations on graphics processing
%units (GPUs) has tended to focus on \emph{regular} data-parallel
%algorithms, for which work to be processed can be evenly split between
%the workgroups that execute a GPU kernel ahead of time.  However,
Many
interesting data-parallel algorithms are \emph{irregular}: the amount
of work to be processed is unknown ahead of time and may be determined
by the computation itself.  There is growing interest in accelerating such algorithms on GPUs, particularly algorithms that
traverse and manipulate linked data structures~\cite{owens-persistent,DBLP:conf/ipps/KaleemVPHP16,DBLP:conf/ipps/DavidsonBGO14,DBLP:conf/hipc/HarishN07,DBLP:journals/topc/MerrillGG15,DBLP:conf/egh/VineetHPN09,DBLP:conf/ppopp/NobariCKB12,DBLP:conf/hpcc/SolomonTT10a,DBLP:conf/popl/PrabhuRMH11,DBLP:conf/ppopp/Mendez-LojoBP12,DBLP:conf/oopsla/PaiP16,DBLP:conf/oopsla/SorensenDBGR16,DBLP:conf/egh/CedermanT08,TPO10,BNP12,Pannotia}.

Irregular algorithms usually require \emph{blocking synchronization}
between workgroups, to balance load and communicate intermediate
results.  For example, many graph algorithms employ a level-by-level strategy, requiring a global barrier between each level, and 
work stealing algorithms require each workgroup
to maintain a queue, which is typically protected by a mutex to enable
stealing by other workgroups.

To operate correctly, a blocking concurrent algorithm requires
\emph{fair} scheduling of workgroups.  Without fairness, a
workgroup may fail to make progress, leading to starvation.  For
example, if one workgroup holds a mutex, an unfair scheduler may cause
another workgroup to spin-wait forever for the mutex to be
released.  Similarly, an unfair scheduler can cause a workgroup to spin-wait
indefinitely at a global barrier so that other workgroups do not reach the barrier.

\vspace{-1mm}
\paragraph{A degree of fairness: occupancy-bound execution}
GPU schedulers are \emph{not} fair. The current GPU
programming models---OpenCL~\cite{opencl2Spec}, CUDA~\cite{cuda-75}  and
HSA~\cite{HSAprogramming11}---specify almost no guarantees regarding scheduling of
workgroups, and current GPUs do not
provide fair scheduling in practice.  Roughly speaking, each workgroup
executing a GPU kernel is mapped to a hardware \emph{compute
  unit}.\footnote{In practice, depending on the kernel, multiple workgroups might map to the same compute unit; we ignore this in our current discussion.}
%
The simplest way for a GPU driver to handle 
the case where more workgroups are launched than there are compute
units is via an \emph{occupancy-bound}
execution model~\cite{owens-persistent,DBLP:conf/oopsla/SorensenDBGR16}.  With this model, once a workgroup has commenced execution on a
compute unit (it has become \emph{occupant}), the workgroup has
exclusive access to the compute unit until it has finished execution.
Experiments suggest that the occupancy-bound execution model is widely
implemented by today's GPU devices and drivers~\cite{owens-persistent,DBLP:conf/oopsla/SorensenDBGR16,DBLP:conf/oopsla/PaiP16,BNP12}.   It is
evidently a simple model for device vendors to implement efficiently.

The occupancy-bound execution model does not guarantee fair scheduling
between workgroups: if all compute units are occupied then a
not-yet-occupant workgroup will not be scheduled until some occupant
workgroup completes execution.  Yet the execution model \emph{does}
provide fair scheduling between \emph{occupant} workgroups, which are
bound to separate compute units that operate in parallel.  Current GPU
implementations of blocking algorithms assume the
occupancy-bound execution model, which they exploit by launching 
no more workgroups than there are available compute
units~\cite{owens-persistent}. This works because \emph{today}'s GPUs seem to provide the
occupancy-bound execution model.

\vspace{-1mm}
\paragraph{Resistance to guaranteeing occupancy-bound execution}

Despite its practical prevalence, none of the current GPU programming
models actually mandate occupancy-bound execution.  Further, there are
clear reasons why this model is undesirable.
% from the point-of-view of
%GPU vendors and end users.
First, the execution model does not enable
multitasking, because a workgroup effectively \emph{owns} a compute
unit until the workgroup has completed execution.  The GPU cannot be used meanwhile for other
tasks, such as graphics rendering.
%Precisely because
%of this problem, operating systems employ a GPU \emph{watchdog} to
%heuristically detect and kill long-running computations on the GPU
%being used for display rendering~\cite{DBLP:conf/iwocl/SorensenD16}.
Second, \emph{energy throttling} is
an important concern for battery-powered devices~\cite{DBLP:journals/comsur/Vallina-RodriguezC13}.  In the future, it will be desirable for a mobile GPU driver to be able to
power down some compute units if the battery level is low or if
the user puts the device into a low power state.  Implementing energy
throttling would require being able to suspend the execution of a
workgroup, power down the compute unit to which the workgroup was
bound, and schedule the workgroup for later execution on a different
compute unit.

Our assessment, informed by discussions with a number of industrial
practitioners who have been involved in the OpenCL and/or HSA
standardisation efforts
(including~\cite{PersonalCommunicationRichards,PersonalCommunicationHowes}
and various engineers from GPU vendors), is that (1) for the above reasons, GPU vendors will
not commit to the occupancy-bound execution model that they currently
implement,
yet (2) GPU vendors will not guarantee fair scheduling, due to the
runtime overhead associated with supporting full preemption of
workgroups.  Vendors instead wish to
retain the essence of the simple occupancy-bound model, supporting preemption
only in key special cases.

\vspace{-1mm}
\paragraph{Our proposal: cooperative kernels}
%
To summarise: blocking algorithms
demand fair scheduling of workgroups, but for good reasons
GPU vendors will not commit even to the guarantees of the
occupancy-bound execution model.

We propose \emph{cooperative kernels}, an extension to the GPU
programming model that aims to resolve this impasse.  A kernel
that requires fair scheduling is identified as \emph{cooperative}, and written using two additional
language primitives, $\offerkill$ and $\offerfork$, placed by the programmer.

At a point where the cooperative kernel could proceed with fewer
workgroups, a workgroup can execute $\offerkill$, offering to
sacrifice itself to the scheduler.  This indicates that the workgroup
would ideally continue executing, but that if required the
scheduler may destroy the workgroup for good.  The cooperative kernel
must be prepared to deal with either scenario.
%
At a point where the cooperative kernel could benefit from having
additional workgroups join the computation, a workgroup can execute
$\offerfork$.  This indicates that the cooperative
kernel is prepared to proceed with the existing set of workgroups, but
is able to take advantage of having one or more additional workgroups
commence execution directly after the $\offerfork$ program point.

The use of $\offerfork$ and $\offerkill$ creates a contract between
the scheduler and (the programmer of the) cooperative kernel.
Functionally, the
scheduler must guarantee that the workgroups executing a cooperative
kernel are fairly scheduled, while the cooperative kernel must be
robust to workgroups leaving and joining the computation in response
to $\offerkill$ and $\offerfork$.  Non-functionally, a cooperative
kernel must ensure that $\offerkill$ is executed frequently enough
that the scheduler can accommodate soft-real time constraints,
e.g.\ allowing a smooth frame-rate for graphics, or allowing compute
units to be powered down quickly when required.  The scheduler should
respond to $\offerkill$ and $\offerfork$ calls in a manner that avoids
under-utilisation of hardware resources by the cooperative kernel,
e.g.\ the scheduler should not kill workgroups unless they are
required for another purpose, and should facilitate forking of
additional workgroups when possible.

The cooperative kernels programming model has several appealing
properties:

\vspace{-1mm}
\begin{enumerate}

\item By providing fair scheduling between workgroups, cooperative
  kernels meet the needs of blocking algorithms, including irregular
  data-parallel algorithms.

\item The model has no impact on the development of regular
  (non-cooperative) compute and graphics kernels: these can be programmed exactly as they
  are now.
%, as can graphics shading routines written using APIs such as
%OpenGL, Vulkan and DirectX.

\item The model is backwards-compatible: if $\offerkill$/$\offerfork$ are ignored, a cooperative kernel will behave
  exactly as a regular kernel does on current GPUs.

\item Cooperative kernels can be implemented over the occupancy-bound
  execution model that current GPUs provide: our proof-of-concept implementation requires
  no additional hardware or driver support.  Thus the model should be eay for vendors to implement directly.

\item Our experiments with a range of applications using three GPUs show that the model can enable efficient multitasking of cooperative and non-cooperative
  tasks.

\end{enumerate}
\vspace{-1mm}

Placing the new primitives requires manual effort, but our experience porting a representative set of
GPU-accelerated irregular algorithms to use cooperative kernels
(\mysec\ref{sec:portingalgorithms}) suggests that this is straightforward in practice.

In summary, our main contributions are:
\emph{cooperative kernels}, an extended GPU programming model that supports the scheduling requirements of blocking algorithms (\mysec\ref{sec:cooperativekernels}); a \emph{proof-of-concept implementation} of cooperative
  kernels on top OpenCL 2.0
  (\mysec\ref{sec:implementation}); \emph{experimental results} assessing the overhead and responsiveness of the cooperative kernels approach over a set of irregular algorithms across three GPUs (\mysec\ref{sec:experiments}).

We begin by providing background on OpenCL via two motivating examples (\mysec\ref{sec:background}).  At the end we discuss related work (\mysec\ref{sec:relatedwork}) and avenues for future work (\mysec\ref{sec:conclusion}).

\section{Background and Motivating Examples}\label{sec:background}

We outline the industry-standard OpenCL programming model on which we
base cooperative kernels (\mysec\ref{sec:opencl}), and illustrate
OpenCL and the scheduling requirements of irregular algorithms using two examples: a work stealing queue and frontier-based graph traversal
(\mysec\ref{sec:openclexamples}).

\subsection{OpenCL Background}\label{sec:opencl}

% \paragraph{OpenCL concepts}
% An OpenCL program is typically executed by many \emph{threads} running
% in parallel. Threads are organized into \emph{workgroups} of equal
% size. The memory is organized in three levels: \emph{global} memory is
% accessible by any thread, \emph{local} memory is shared at the workgroup
% level (i.e., threads of different workgroups cannot access the same
% local memory), and \emph{private} memory is dedicated to thread-level
% storage and cannot be shared between threads. These concepts are
% illustrated over the two examples.

An OpenCL program is divided into \emph{host} and \emph{device}
components.  A host application runs on the CPU and launches one or
more \emph{kernels} that run on accelerator devices---GPUs in the
context of this paper.  A kernel is written in OpenCL C, based on C99.
All threads executing a kernel execute the same entry function with
identical arguments.  A thread can call $\getglobalid()$
to obtain its unique id, to access distinct data or follow different control flow paths to other threads.

The threads executing a kernel are divided into a number of \emph{workgroups}, each consisting of a fixed number of threads.  Builtin functions
$\getlocalid()$ and $\getgroupid()$ return a thread's local id within
its workgroup and the id of the workgroup, respectively.\footnote{For simplicity we assume 1D kernels, which captures all the
irregular algorithms we know of, using
e.g.\ $\getglobalid()$ to mean $\getglobalid(0)$.
}  The number
of threads per workgroup and the number of workgroups are obtained via
$\getlocalsize()$ and $\getnumgroups()$.  A thread's local and global
id are related via the equation $\getglobalid() = \getgroupid() \times
\getlocalsize() + \getlocalid()$.  The total number of threads executing the kernel is given by $\getglobalsize() = \getlocalsize()\times\getnumgroups()$.

Execution of the threads in a workgroup can be synchronised via a
workgroup barrier, which also ensures memory consistency.  This
barrier is a \emph{workgroup-level} function: it can only
appear in conditional code if all threads in a workgroup that reach
the barrier agree on the guards of all enclosing conditional
statements.  A \emph{global} barrier
(synchronising all threads of a kernel) is \emph{not} provided as
a primitive.

\paragraph{Memory spaces and memory model}
An OpenCL kernel has access to four memory spaces.  \emph{Shared virtual
memory} (SVM) is accessible to all threads and the host application concurrently.  \emph{Global} memory is shared among all threads executing a
kernel.  Each workgroup has its own portion of \emph{local} memory for fast intra-workgroup communication.  Every thread has
an allocation of very fast \emph{private} memory, where
function-local variables are allocated.

Communication within a workgroup can be achieved
using a workgroup barrier.  Finer-grained
communication within a workgroup, as well as inter-workgroup
communication and communication with the host while the kernel is
running, is enabled by a set of atomic data types and operations.  In
particular, fine-grained host/device communication is via atomic
operations on SVM.

\paragraph{Execution model}

OpenCL specifically makes no guarantees about fair scheduling between
workgroups executing the same kernel, stating~\cite[p.\ 31]{opencl2Spec}: \emph{``A
  conforming implementation may choose to serialize the workgroups 
%so
%  a correct algorithm cannot assume that workgroups will execute in
%  parallel.  
\dots There is no safe and portable way to synchronize across
  the independent execution of workgroups since once in the work-pool,
  they can execute in any order.''}  CUDA similarly provides no guarantees~\cite{cuda-75}.
%
HSA provides limited, one-way guarantees,
stating~\cite[p. 46]{HSAprogramming11}: \emph{Work-group A can wait
  for values written by work-group B without deadlock provided ... A
  comes after B in work-group flattened ID order}. This is not sufficient to support blocking algorithms that use
mutexes and inter-workgroup barriers, both of which require \emph{symmetric} communication between
threads.

\subsection{Motivating Examples}\label{sec:openclexamples}

\begin{figure}

\begin{lstlisting}
(*@\label{line:wksteal:kernelfunc}@*)kernel work_stealing(global Task * queues) {
(*@\label{line:wksteal:getgroupid}@*)  int queue_id = get_group_id();
(*@\label{line:wksteal:mainloop}@*)  while (more_work(queues)) {
(*@\label{line:wksteal:poporsteal}@*)    Task * t = pop_or_steal(queues, queue_id);
    if (t) {
      // task processing may create more work
(*@\label{line:wksteal:processtask}@*)      process_task(t, queues, queue_id);
    }
  }
}
\end{lstlisting}

\caption{An excerpt of a work stealing algorithm in OpenCL}\label{fig:workstealing}
\end{figure}

\begin{figure}

\begin{lstlisting}
kernel graph_app(global graph * g, 
                 global nodes * n0, global nodes * n1) {
  int level = 0;
  global nodes * in_nodes = n0;
  global nodes * out_nodes = n1;
  int tid = get_global_id();
  int stride = get_global_size();
(*@\label{line:graph:iterate}@*)  while(in_nodes.size > 0) {
    for (int i = tid; i < nodes.size; i += stride) {
      process_node(g, in_nodes[i], out_nodes, level);
    }
(*@\label{line:graph:swap}@*)    swap(&in_nodes, &out_nodes);
(*@\label{line:graph:gb1}@*)    global_barrier();
(*@\label{line:graph:reset}@*)    reset(out_nodes);
    level++;
(*@\label{line:graph:gb2}@*)    global_barrier();
  }
}
\end{lstlisting}
\caption{An OpenCL graph traversal algorithm, using a global barrier for synchronisation}\label{fig:graphsearch}
\end{figure}

\paragraph{Work stealing example}
%
Work stealing enables dynamic balancing of tasks across several
processing units. It is useful when the number of tasks to be
processed is dynamic, due to one task creating an arbitrary number of new tasks.  In the context
of GPUs, each workgroup has an associated queue from which it
obtains tasks to process, and to which it stores newly created
tasks. If a workgroup's queue is empty, the workgroup tries to \emph{steal} a task from an other
queue. Previous work addresses the feasibility and efficiency of work
stealing on GPUs~\cite{DBLP:conf/egh/CedermanT08,TPO10}.

\myfiglong\ref{fig:workstealing} shows a simplified version of an OpenCL
kernel that uses a work stealing queue. Each thread receives a pointer to the task
queues, in global memory, initialized by the
host to contain the initial tasks. Each thread identifies
the workgroup it belongs to via $\getgroupid()$
(line~\ref{line:wksteal:getgroupid}), which is used as a queue id and used to access the relevant task queue. The
$\mathsf{pop\_or\_steal}$ function (line~\ref{line:wksteal:poporsteal})
is called by all threads of a workgroup.  Inside
$\mathsf{pop\_or\_steal}$ only the thread with a local id equal to $0$,
the \emph{master thread}, will try to pop a
task from the workgroup's queue, or steal a task from a different
queue. The master thread uses local memory inside
$\mathsf{pop\_or\_steal}$ to ensure that the returned value is valid for
all threads of its workgroup. If a task was obtained, then all threads
of the workgroup participate in the processing of the task (line~\ref{line:wksteal:processtask}).
This may lead to the creation of further tasks, which
the master thread will push to the workgroup's queue.

Although not depicted here, concurrent accesses to queues inside
$\mathsf{more\_work}$ and $\mathsf{pop\_or\_steal}$ are guarded by a
mutex per queue, implemented using atomic compare and swap operations
on global memory.  The kernel presents two opportunities
for spin-waiting: spinning to obtain a mutex, and spinning in the main
kernel loop to obtain a task.  The algorithm thus depends
on a fair scheduler, which is not guaranteed by current GPU programming models.

\paragraph{Graph traversal example} \myfiglong\ref{fig:graphsearch} shows the
skeleton of a frontier-based graph traversal algorithm; such algorithms have
been shown to be execute efficiently on GPUs~\cite{BNP12,DBLP:conf/oopsla/PaiP16}. 
The kernel is
given three arguments in global memory: a graph structure, and two
arrays that can contain graph nodes. Initially, $\keyword{n0}$ contains the
starting nodes to process. Private variable $\keyword{level}$ indicates the frontier level currently being
processed, and $\keyword{in\_nodes}$ and $\keyword{out\_nodes}$ point to
distinct node arrays recording the nodes to be processed during the current and next frontier, respectively.

The application iterates as long as the current frontier contains
nodes to process (line~\ref{line:graph:iterate}). At each frontier,
the nodes to be processed are evenly distributed between the executing
threads. This is done through \emph{stride} based processing. 
%
In this case, the stride is the total number of threads, obtained via 
$\getglobalsize()$.  Each thread processes nodes with the following sequence of indices:
\code{$\keyword{tid}$+$\keyword{stride}$*0},
\code{$\keyword{tid}$+$\keyword{stride}$*1},
\code{$\keyword{tid}$+$\keyword{stride}$*2}, etc., where $\keyword{tid}$ is a thread's global id.
A thread processes a node via $\keyword{process\_node}$, which may (a) push nodes to process in the next frontier to
$\keyword{out\_nodes}$ and (b) use the current frontier, $\keyword{level}$, in
the computation. After processing the frontier, the threads swap their
node array pointers (line~\ref{line:graph:swap}).

At this point, the GPU threads must wait for all other threads to
finishing processing the frontier. To achieve
this, we use a global barrier construct
(line~\ref{line:graph:gb1}). After all threads reach this point, the
output node array is reset (line~\ref{line:graph:reset}) and the level
is incremented. The threads use another global barrier to wait until the output node is
reset (line~\ref{line:graph:gb2}), after which the threads can continue to the next frontier.

The global barrier used in this application is not provided as a GPU
primitive, though previous works have shown that such a global
barrier can be implemented~\cite{XF10,DBLP:conf/oopsla/SorensenDBGR16},
based on CPU barrier designs~\cite[ch. 17]{HS08}.
 These barriers employ spinning to ensure threads wait at the barrier until all
threads have arrived, which requires fair
scheduling between workgroups for the barrier to operate correctly.

The mutexes and barriers used by these two examples appear to run
reliably on current GPUs for kernels that are executed with no more
workgroups than there are compute units.  This is due to the fairness
of the occupancy-bound execution model that current GPUs have been
shown, experimentally, to provide.  But, as discussed in
\mysec\ref{sec:intro}, this model is not endorsed
language standards or vendor implementations, and is unlikely to be respected in the future.
%
In \mysec\ref{sec:programmingguidelines} we show how the work stealing
and graph traversal examples of \myfigs\ref{fig:workstealing} and~\ref{fig:graphsearch} can be
updated to use our cooperative kernels programming model to resolve
the scheduling issue.

\section{Cooperative Kernels}\label{sec:cooperativekernels}

We present our cooperative kernels programming model as an extension
to OpenCL.  However, the cooperative kernels concept is more general,
and could be applied to extend other GPU programming models, such as
CUDA and HSA.  We describe the semantics of the model
(\mysec\ref{sec:semantics}), use our motivating examples to discuss
programmability (\mysec\ref{sec:programmingguidelines}), outline
important nonfunctional properties that the model requires to work
successfully (\mysec\ref{sec:nonfunctional}) and explain that the
model is backwards-compatible
(\mysec\ref{sec:backwardscompatibility}).
%
In Appendices~\ref{appendix:semantics} and~\ref{appendix:semanticalternatives} (supplementary material)
we present a formal semantics for cooperative kernels atop an abstract GPU programming model,
and discuss several cases where we
might have taken different and also reasonable semantic decisions.

\subsection{Semantics of Cooperative Kernels}\label{sec:semantics}

As with a regular OpenCL kernel (see \mysec\ref{sec:opencl}), a
cooperative kernel is launched by the host application.  The host
application passes parameters to the kernel and specifies a desired
number of workgroups, each consisting of a specified number of
threads.  Unlike in a regular kernel, the parameters to a cooperative kernel are immutable (though pointer
parameters can refer to mutable data).

Cooperative kernels are written using the following 
extensions: $\transmit$, a qualifier on the variables of a
thread; $\offerkill$ and $\offerfork$, the key functions that enable
cooperative scheduling; and $\globalbarrier$ and $\resizingglobalbarrier$
primitives for inter-workgroup synchronisation.
%
We present the semantics of these constructs intuitively, using English.
In Appendix~\ref{appendix:semantics} (see supplementary material) we present an abstract operational semantics for our language extensions in a simple GPU-like programming model.

\paragraph{Transmitted variables}

A variable declared in the root scope of the cooperative kernel can
optionally be annotated with a new $\transmit$ qualifier.  Annotating
a variable $v$ with $\transmit$ means that when a workgroup spawns new
workgroups by calling $\offerfork$, the workgroup should transmit its
current value for $v$ to the threads of the new workgroups, to serve
as an initial value for $v$.  We detail the semantics for this when we
describe $\offerfork$ below.

\paragraph{Active workroups}

If the host application launches a cooperative kernel requesting $N$
workgroups, this indicates that the kernel should be executed with a
maximum of $N$ workgroups, and that as many workgroups as possible, up
to this limit, are desired.  However, the scheduler may initially
schedule fewer than $N$ workgroups, and as explained below the number
of workgroups that execute the cooperative kernel can change during
the lifetime of the kernel.

The number of \emph{active workgroups}---workgroups executing the
kernel---is denoted $M$.  The active workgroups
have consecutive workgroup ids in the range $[0, M-1]$.
Initially, at least one workgroup is active;
if necessary the scheduler must postpone execution of the kernel until
a some compute unit becomes available.

When executed by a cooperative
kernel, $\getnumgroups()$ returns $M$, the \emph{current} number of
active workgroups.  This is in contrast to $\getnumgroups()$ for
regular kernels, which returns the fixed number of workgroups that
were requested at kernel launch time (see \mysec\ref{sec:opencl}).

Fair scheduling \emph{is} guaranteed between active workgroups.  That is, if
some thread in an active workgroup is enabled then eventually some
thread in the active workgroup is guaranteed to execute an
instruction.

%Note that our programming model does not mandate that
%threads within a workgroup must be fairly scheduled, and indeed this
%is often not the case in GPU programming models, because threads are
%often organised into \emph{warps} or \emph{wavefronts} within a
%workgroup, exhibiting lock-step predicated execution that is
%inherently unfair.

\paragraph{Semantics for $\offerkill$}

The $\offerkill$ primitive allows the cooperative kernel to return
compute units to the scheduler by offering to sacrifice workgroups.
The idea is as follows: allowing the scheduler to arbitrarily and abruptly terminate execution
of workgroups might be drastic, yet the kernel
may contain specific program points at which a workgroup could
\emph{gracefully} leave the computation.  We give examples of these
for work stealing and graph traversal in \mysec\ref{sec:programmingguidelines}.

Similar to the OpenCL workgroup $\mathsf{barrier}$ primitive,
$\offerkill$, is a workgroup-level function---it must be encountered
uniformly by all threads in a workgroup (see
\mysec\ref{sec:opencl}).

Suppose a workgroup with id $m$ executes $\offerkill$.  
If the workgroup has the highest largest id of all active workgroups then it can be killed by the scheduler, except that workgroup 0 can never be killed (to ensure that the kernel computation does not terminate early).  More formally, if $m < M-1$
or $M=1$ then $\offerkill$ is a no-op.  If instead $M > 1$ and $m =
M-1$, the scheduler can choose to ignore the offer, so that $\offerkill$
executes as a no-op, or accept the offer, so that execution of the workgroup ceases and
the number of active workgroups $M$ is atomically decremented by one.  In the second case, the scheduler can assign some other task to
the compute unit on which the workgroup was executing.

\paragraph{Semantics for $\offerfork$}

Recall that
a desired limit of $N$ workgroups was specified when the cooperative kernel was launched, but that the number of active workgroups, $M$, may be smaller
than $N$, either because (due to competing workloads) the scheduler
did not provide $N$ workgroups initially, or because the kernel has
given up some workgroups via $\offerkill$ calls.  Through the
$\offerfork$ primitive (also a workgroup-level function), the kernel and scheduler can collaborate to allow new
workgroups to join the computation at an appropriate point and with
appropriate state.

Suppose a workgroup with id $m\leq M$ executes $\offerfork$.  Then the following occurs: an
integer $k \in [0, N-M]$ is chosen by the scheduler;
$k$ new workgroups are spawned with consecutive ids in the range $[M,
  M+k-1]$; the active workgroup count $M$ is atomically incremented by $k$.

The $k$ new workgroups commence execution at the program point
immediately following the $\offerfork$ call.  The variables that
describe the state of a thread are all uninitialised for the threads
in the new workgroups; reading from these variables without first
initialising them is an undefined behaviour.  There are two exceptions
to this: (1) because the parameters to a cooperative kernel are
immutable, the new threads have access to these parameters as part of
their local state and can safely read from them; (2) for each variable
$v$ annotated with $\transmit$, every new thread's copy of $v$ is
initialised to the value that thread 0 in workgroup $m$ held for $v$
at the point of the $\offerfork$ call.
%
In effect, thread 0 of the forking workgroup \emph{transmits} the relevant
portion of its local state to the threads of the forked workgroups.

%We discuss in \mysec\ref{sec:semanticalternatives} the rationale for
%having all new threads obtain this local state portion from thread 0,
%as well as our reasons for selecting which variables to transmit via
%annotations rather than transmitting the entire thread state.

Notice that $k=0$ is always a legitimate choice for the number of
workgroups to be spawned by an $\offerfork$ call, and if the number of
active workgroups $M$ is equal to the workgroup limit $N$, $k=0$ is
guaranteed.

\paragraph{Global barriers}

Many irregular algorithms require synchronisation across workgroups.
This can be achieved via a \emph{global barrier}, as illustrated in the graph traversal example of \mysec\ref{sec:openclexamples} and \myfig\ref{fig:graphsearch}.
Because a global barrier requires fair scheduling of workgroups it cannot be safely implemented in OpenCL, CUDA or HSA, and thus is not provided as a primitive in any of these programming models.

Workgroups of a \emph{cooperative} kernel \emph{are} fairly scheduled, so a
global barrier primitive can be provided, and we specify two variants: $\globalbarrier$
and $\resizingglobalbarrier$.

Our $\globalbarrier$ primitive is a kernel-level function: if it
appears in conditional code then it must be reached by \emph{all}
threads executing the cooperative kernel.  On reaching a
$\globalbarrier()$, a thread waits until all threads have arrived at
the barrier.  Once all threads have arrived, the threads may proceed
past the barrier with the guarantee that all global memory accesses
issued before the barrier have completed.  The $\globalbarrier$
primitive can be implemented by adapting an inter-workgroup barrier
design, e.g.~\cite{XF10}, to take account of a growing and shrinking number of workgroups, and the atomic operations provided by
the OpenCL 2.0 memory model enable a memory-safe
implementation~\cite{DBLP:conf/oopsla/SorensenDBGR16}.  However, implementing such a barrier is
involved, hence why we include this function as a primitive.

The $\resizingglobalbarrier$ primitive is also a kernel-level
function.  It is identical to $\globalbarrier$, except that it caters
for cooperation with the scheduler: by issuing a
$\resizingglobalbarrier$ the programmer indicates that the cooperative
kernel is prepared to proceed after the barrier with fewer workgroups,
or with additional workgroups joining the computation.

When all threads have reached $\resizingglobalbarrier$,
the number of active workgroups, $M$, is atomically set to a new value, $M'$ say, with $0 < M' \leq N$.
If $M' = M$ then the active workgroups remain unchanged.  If $M' < M$, workgroups $[M', M-1]$ are
killed.  If $M' > M$ then $M'-M$ new workgroups join the computation after the barrier,
as if they were forked from workgroup 0.  In particular, the
$\transmit$-annotated local state of thread 0 in workgroup 0 is
transmitted to the threads of the new workgroups.

The semantics of $\resizingglobalbarrier$ can be modelled via the following sequence of calls:

\lstset{basicstyle=\tt,numbers=none}
\begin{lstlisting}
  $\globalbarrier()$;
  if($\getgroupid()$ == 0) $\offerfork$();
  $\globalbarrier()$;
  $\offerkill()$;
  $\globalbarrier()$;
\end{lstlisting}
\lstset{basicstyle=\scriptsize\tt,numbers=left}

The enclosing $\globalbarrier$ calls ensure that the change in number
of active workgroups from $M$ to $M'$ occurs entirely within the
resizing barrier, so that from a programmer's perspective, $M$ changes atomically.  The middle $\globalbarrier$ ensures that forking occurs before killing, so that while the barrier may kill a workgroup with id $i$, it should not also fork a new workgroup with id $i$.  That is, workgroups $[0, \textrm{min}(M, M') - 1]$ should be left intact.

Because $\resizingglobalbarrier$ can be implemented in the above
manner, we do not regard it \emph{conceptually} as a primitive of our
programming model.  However, we discuss in
\mysec\ref{sec:resizingbarrier} how $\resizingglobalbarrier$ can be
implemented significantly more efficiently if it is treated as a
primitive operation that interacts directly with the scheduler.

\subsection{Programming With Cooperative Kernels}\label{sec:programmingguidelines}

We discuss the main semantic change induced by our programming
model---that the number of workgroups executing a kernel varies
dynamically---and illustrate the new language constructs and the
impact of this semantic change by adapting the work stealing and graph
traversal algorithms of \mysec\ref{sec:openclexamples} to use
cooperative kernels.

\paragraph{A changing number of workgroups}  Unlike in regular OpenCL,
the value returned by $\getnumgroups()$ is not fixed during the
lifetime of a cooperative kernel: it corresponds to the active group count $M$ and may change due to workgroups executing
$\offerkill$, $\offerfork$ and $\resizingglobalbarrier$.  Being a function of $\getnumgroups()$, the value returned by $\getglobalsize()$
is also subject to change.
A cooperative kernel must therefore be written in a manner that is
robust to changes in the values returned by these functions.

In general, their volatility means that use of these functions should be avoided.
However, the situation is more stable if a cooperative kernel does not directly
issue $\offerkill$ and $\offerfork$ calls directly.  In this case,
only calls to $\resizingglobalbarrier$ can cause the number of active
workgroups to change.  At any point
during execution, the threads of a kernel are executing between some
pair of resizing barrier calls, which we call a \emph{resizing barrier
  interval} (here we consider the kernel entry and exit points
conceptually to be special cases of resizing barriers).  The number of
workgroups executing the kernel is thus constant within each resizing
barrier interval, i.e.\ the result returned by repeated calls to
$\getnumgroups()$ will be identical during such an interval; the same holds for $\getglobalsize()$.
%
This guarantee can be exploited by algorithms that perform strided
data processing, which we illustrate below using the graph traversal
example.

\paragraph{Adapting work stealing}

In this example there is no state to transmit since a computation is
entirely parameterised by a task, which is retrieved from a queue
located in global memory. With respect to \myfig~\ref{fig:workstealing},
we add $\offerfork$ and $\offerkill$ calls at the start of the main loop
(below line~\ref{line:wksteal:mainloop}) to let a workgroup offer itself
to be killed or forked, respectively, before it processes a task.  Note
that a workgroup may be killed even if its associated task queue is not
empty, since remaining tasks will be stolen by other workgroups. In
addition, since $\offerfork$ may be the entry point of a workgroup, the
queue id must now be computed after it, so we move
line~\ref{line:wksteal:getgroupid} to be placed just before
line~\ref{line:wksteal:poporsteal}. In particular, the queue id cannot
be transmitted since we want a newly spawned workgroup to read its own
queue and not the one of the forking workgroup.

% \begin{figure}
% \begin{lstlisting}
% kernel work_stealing(global Task * queues) {
% (*@\label{line:cwksteal:morework}@*)  while (more_work(queues)) {
% (*@\label{line:cwksteal:offerfork}@*)    offer_fork();
% (*@\label{line:cwksteal:offerkill}@*)    offer_kill();
% (*@\label{line:cwksteal:getgroupid}@*)    int queue_id = get_group_id();
% (*@\label{line:cwksteal:poporsteal}@*)    Task * t = pop_or_steal(queues, queue_id);
%     if (t) {
%       // task processing may create more work
% (*@\label{line:cwksteal:processtask}@*)      process_task(t, queues, queue_id);
%     }
%   }
% }
% \end{lstlisting}
% \caption{Cooperative version of the work stealing kernel of
%   \myfig~\ref{fig:workstealing}, using $\offerfork$ and
%   $\offerkill$}\label{fig:workstealing-cooperative}
% \end{figure}

\paragraph{Adapting graph traversal} 
\myfiglong\ref{fig:cgraphsearch} shows a cooperative version of the
graph traversal kernel of \myfig\ref{fig:graphsearch} from
\mysec\ref{sec:openclexamples}.  On lines~\ref{line:cgraph:resizing1}
and ~\ref{line:cgraph:resizing2}, we change the original global
barriers into a resizing barriers. Several variables are marked to be
transmitted in the case of workgroups joining at the resizing barriers
(lines~\ref{line:cgraph:transmit1}, \ref{line:cgraph:transmit2} and
\ref{line:cgraph:transmit3}): $\keyword{level}$ must be restored so
that new workgroups know which frontier they are processing;
$\keyword{in\_nodes}$ and $\keyword{out\_nodes}$ must be restored so
that new workgroups know which of the node arrays to use for input and
output. Lastly, the static work distribution of the original kernel is
no longer valid in a cooperative kernel. This is because the stride
(which is based on $M$) may change after each resizing barrier
call. To fix this, we re-distribute the work after each resizing
barrier call by recomputing the thread id and stride
(lines~\ref{line:cgraph:rechunking1} and
\ref{line:cgraph:rechunking2}). This example exploits the fact that
the cooperative kernel does not issue $\offerkill$ nor $\offerfork$
directly: the value of $\keyword{stride}$ obtained from
$\getglobalsize()$ at line~\ref{line:cgraph:rechunking2} is stable
until the next resizing barrier at line~\ref{line:cgraph:resizing1},
and in particular can be safely used inside the $\keyword{for}$ loop.

\begin{figure}

\begin{lstlisting}
kernel graph_app(global graph *g, 
                 global nodes *n0, global nodes *n1) {
(*@\label{line:cgraph:transmit1}@*)  transmit int level = 0;
(*@\label{line:cgraph:transmit2}@*)  transmit global nodes *in_nodes = n0;
(*@\label{line:cgraph:transmit3}@*)  transmit global nodes *out_nodes = n1;
  while(in_nodes.size > 0) {
(*@\label{line:cgraph:rechunking1}@*)    int tid = get_global_id();
(*@\label{line:cgraph:rechunking2}@*)    int stride = get_global_size();
    for (int i = tid; i < nodes.size; i += stride) {
      process_node(g, in_nodes[i], out_nodes, level);
    }
    swap(&in_nodes, &out_nodes);
(*@\label{line:cgraph:resizing1}@*)    resizing_global_barrier();
    reset(out_nodes);
    level++;
(*@\label{line:cgraph:resizing2}@*)    resizing_global_barrier();
  }
}
\end{lstlisting}
\caption{Cooperative version of the graph traversal kernel of \myfig\ref{fig:graphsearch}, using a resizing global barrier and $\transmit$ annotations}\label{fig:cgraphsearch}
\end{figure}

\paragraph{Patterns for irregular algorithms}
In \mysec\ref{sec:portingalgorithms} we describe the set of irregular GPU algorithms used
in our experiments, which largely captures the irregular blocking
algorithms that are available as open source GPU kernels.  These all
employ either work stealing or operate on graph data structures, and placing our new constructs follows a common, easy-to-follow pattern in each case.

Like the simplified example of \myfig\ref{fig:workstealing-cooperative}, the work stealing algorithms have a transactional flavour
and require little or no state to be carried between transactions.  Thus the point at which a workgroup is ready to process a new task is a natural point to place $\offerkill$ and $\offerfork$, and few or no $\transmit$ annotations are required.

The example of \myfig\ref{fig:cgraphsearch} is simpler than, but representative of,
most level-by-level graph algorithms.
It is typically the case that on completing a level of
the graph algorithm, the next level could be processed by more or
fewer workgroups, something which $\resizingglobalbarrier$
facilitates.  As in \myfig\ref{fig:cgraphsearch}, some level-specific state must be transmitted to new workgroups.

Our practical experience suggests that cooperative kernels are a good
fit for at least these two classes of irregular algorithms, which is
of course partly because we were inspired by these algorithms in
designing the model.

\subsection{Non-Functional Requirements}\label{sec:nonfunctional}

The semantics presented in \mysec\ref{sec:semantics} describe the envelope of
behaviour that a developer of a cooperative kernel should be prepared
for.
%
However, the aim of cooperative kernels is to find a balance that
allows \emph{efficient} execution of algorithms that require fair scheduling, and
\emph{responsive} multitasking, so that the GPU can be shared between
cooperative kernels and other shorter tasks with soft real-time constraints.
%
To achieve this balance, an implementation of the cooperative
kernels model, and the programmer of a cooperative kernel, must strive
to meet the following non-functional requirements.

\paragraph{Sufficient $\offerkill$ calls}

The purpose of $\offerkill$ is to provide the scheduler with an
opportunity to destroy a workgroup in order to schedule
higher-priority tasks.  The scheduler relies on the cooperative kernel
to execute $\offerkill$ sufficiently frequently that soft real-time
constraints of other workloads can be met.
%
Using our work stealing example: in
\myfig\ref{fig:workstealing-cooperative} a workgroup offers itself to
the scheduler after processing each task.  If tasks are sufficiently
fast to process then the scheduler will have ample opportunities to
de-schedule workgroups.  But if tasks can be very time-consuming to
process then it might be necessary to rewrite the algorithm so that
tasks are shorter and more numerous, to achieve a higher rate of calls
to $\offerkill$.

Getting this non-functional requirement right depends on the
constraints of the tasks to be co-scheduled with a cooperative kernel,
on properties of the cooperative kernel itself, and on the efficiency
of the GPU that is used for execution.  In \mysec\ref{sec:sizingnoncoop} we conduct
experiments to understand the response rate that would be required to
co-schedule graphics rendering with a cooperative kernel in order to
achieve a smooth frame rate, and we evaluate responsiveness in
practice in \mysec\ref{sec:responsiveness}.

\paragraph{Sufficient workgroups for the cooperative kernel}

Recall that the cooperative kernel would ideally utilise the $N$
workgroups requested upon kernel launch, in order to conduct its
processing efficiently.  The scheduler should thus aim to provide the
cooperative kernel $N$ workgroups if other constraints allow it,
by accepting an $\offerkill$ only if a compute unit is required for another
task, and responding positively to $\offerfork$ calls if compute units are available.  Our
implementation of cooperative scheduling (\mysec\ref{sec:implementation}) is as
generous as possible to the cooperative kernel in this regard.

Note that the above guidelines are not hard-and-fast rules.  For
example, a smart scheduler might respond selectively to $\offerkill$
and $\offerfork$ calls by predicting future high-priority workloads
based on past workloads.

\subsection{Backwards Compatibility}\label{sec:backwardscompatibility}

A goal of coopeative kernels is to allow traditional, non-blocking GPU
tasks (e.g.\ graphics shaders in OpenGL, Vulkan and DirectX, and regular data-parallel computational kernels)
to be co-scheduled with blocking irregular algorithms.  It is
important to note that our new model has no effect on the manner in
which these traditional GPU workloads are programmed.

Furthermore, if our cooperative language extensions are
ignored---$\offerkill$ and $\offerfork$ are treated as no-ops,
$\transmit$ annotations are ignored, and $\globalbarrier$ and
$\resizingglobalbarrier$ calls are redirected to an existing global
barrier implementation (so that no resizing occurs in the latter
case)---then a cooperative kernel will behave just like its
non-cooperative counterpart would behave.  As a result, a developer
for whom today's occupancy-bound execution model suffices can upgrade
their kernel to a cooperative form and retain backwards-compatibility
with today's GPU platforms by pre-processing away our language
extensions.

\section{Prototype Implementation}\label{sec:implementation}

Our vision is that support for cooperative kernels will be integrated
in the runtimes of future GPU implementations of OpenCL and related
programming models, with driver and compiler support for our new
primitives.  To experiment with our ideas on current GPUs, we have
developed a prototype that mocks up the required runtime support via a
\emph{megakernel}, and exploits the occupancy-bound execution model
that these GPUs provide to ensure fair scheduling between workgroups.
We emphasise that an aim of cooperative kernels is to \emph{avoid}
depending on the occupancy-bound model.  Our protoytpe exploits this
model simply to allow us to experiment with current GPUs whose
proprietary drivers we cannot change.  We describe the megakernel approach (\mysec\ref{sec:megakernel})
and detail various aspects of the scheduler component of our implementation (\mysec\ref{sec:schedulerimpl}).

\subsection{The megakernel mock up}\label{sec:megakernel}

Instead of multitasking multiple separate kernels, we merge a set of
kernels into a megakernel---a single, monolithic kernel.  The
megakernel is launched with as many workgroups as can be occupant
concurrently.  One workgroup takes the role of scheduler, and the
scheduling logic is embedded as part of the megakernel.  The remaining
workgroups act as a pool of workers.  A worker repeatedly queries the
scheduler to be assigned a task.  A task corresponds to executing a
cooperative or non-cooperative kernel.  In the non-cooperative case,
the workgroup executes the relevant kernel function uninterrupted,
then awaits further work.  In the cooperative case, the workgroup
either starts from the kernel entry point or immediately jumps to a
designated point within the kernel, depending on whether the workgroup
is an initial workgroup of the kernel, or a forked workgroup.  In the
latter case, the new workgroup also receives a struct containing the
values of all relevant $\transmit$-annotated variables.

\paragraph{Simplifying assumptions}

For ease of implementation, our current prototype supports
multitasking a single cooperative kernel with a single non-cooperative
kernel (though the non-cooperative kernel can be invoked many times).
Rather than adding language support for $\transmit$ annotations, we
assume that all variables declared at the root lexical scope of a
cooperative kernel are $\transmit$-annotated; this allows our
implementation to piggy-back on top of Clang (see below) without
requiring front-end changes.  Finally, we require that $\offerkill$,
$\offerfork$ and $\resizingglobalbarrier$ are called from the entry
function of a cooperative kernel.  This allows us to use
$\keyword{goto}$ and $\keyword{return}$ to direct threads into and out
of the kernel.  With these restrictions we can experiment
with interesting irregular algorithms (see
\mysec\ref{sec:experiments}).  A non-mock implementation of
coopeartive kernels would not use the megakernel approach, so we do not deem
the engineering effort associated with lifting these restrictions to
be worthwhile.

\paragraph{Merging kernels}

We have built a tool, \kernelmerge{}, that takes a cooperative kernel
and a non-cooperative kernel and merges them to produce a megakernel
with an embedded GPU scheduler.  We implemented \kernelmerge{} using the Clang LibTooling framework,\footnote{\url{http://clang.llvm.org/docs/LibTooling.html}}
leveraging recent Clang support for OpenCL 2.0~\cite{ClangOpenCL20}.

\subsection{Scheduler design}\label{sec:resizingbarrier}\label{sec:schedulerimpl}

To enable multitasking through cooperative kernels, the runtime (in
our case, the mega-kernel) must track the state of all
workgroups, i.e.\ whether a workgroup is waiting or computing a kernel,
maintain consistent context states for each kernel, e.g.\ tracking the number
of active workgroups, and provide a
safe way for these states to be modified in response to $\offerfork$/$\offerkill$. We discuss these issues,
and describe the implementation of an efficient resizing barrier.

We describe how the scheduler would handle arbitrary
combinations of kernels, though as noted above our current
implementation is restricted to the case of two kernels.

\paragraph{Kernel and scheduler contexts}

When a workgroup executes a kernel (cooperative or otherwise), it must
be able to query a consistent context to obtain a workgroup id value
and the
number of workgroups executing the kernel. For multi-kernel
kernel execution, we emulate the $\getnumgroups$ and $\getgroupid$ 
primitives via a per kernel \emph{context}.  For a given kernel this records: the number of active groups $M$ (irrelevant for a non-cooperative kernel), the total number of requested groups $N$, and a mapping from the native workgroup id to a new workgroup
id valid for the kernel.

To dynamically manage workgroups executing cooperative
kernels, our framework must track the state of each workgroup and
provide a channel of communication from the scheduler workgroup to
workgroups executing $\offerfork$ and $\offerkill$. To achieve this,
we use a \emph{scheduler context} structure, mapping a primitive workgroup id the workgroup's status, which is either \emph{available} or the id of the kernel that the workgroup is currently executing.  The scheduler can then send cooperative
kernels a \emph{resource message}, commanding workgroups to exit at $\offerkill$, or 
spawn additional workgroups at $\offerfork$. Thus, the
scheduler context needs a communication channel for each cooperative
kernel. We implement the communication channels using atomic variables in global memory.

\paragraph{Launching kernels and managing workgroups}

To launch a kernel, the host sends a data packet to the GPU
scheduler consisting of: a kernel to execute, the inputs to the kernel, the
number of workgroups that should execute the kernel, and a flag indicating whether the kernel
is cooperative. In our implementation, this host-device communication channel is built using fine-grained SVM atomics.

On receiving a data packet
describing a kernel launch $K$, the GPU scheduler must make decisions
about how to schedule $K$. Suppose $K$ requests
$N$ workgroups. The scheduler queries the scheduler context.  If
there are at least $N$ available workgroups, $K$ can be scheduled
immediately. Suppose instead that there are only $N_a < N$ available workgroups, but a cooperative kernel $K_c$ is
executing. The scheduler can use $K_c$'s channel in the scheduler context to command $K_c$ to provide $N - N_a$ workgroups via $\offerkill$.
Once $N$ workgroups are available, the scheduler initialises a
kernel context $\KC$ for $K$, setting the number of active workgroups
to $N$ and mapping available workgroups (located in the scheduler
context) to their new $\KC$ workgroup ids. The scheduler then sends $N$
workgroups from the available workgroups to execute kernel $K$ with
kernel context $\KC$.  (Note that if the new kernel $K$ is itself a cooperative kernel, the scheduler would be free to provide $K$ with fewer than $N$ active workgroups initially.)

If a cooperative kernel $K_c$ is executing with fewer workgroups than it initially requested, the scheduler may decide make extra workgroups available
to $K_c$, to be obtained next time $K_c$ calls $\offerfork$.  To do this, the scheduler asynchronously signals $K_c$ through $K_c$'s
channel to indicate the number of
workgroups that should join at the next $\offerfork$ command.
When a workgroup $w$ of $K_c$ subsequently executes $\offerfork$, thread 0 of $w$ updates the kernel and scheduler contexts so that the given number of new workgroups are directed to the program point after the $\offerfork$ call.  This involves selecting workgroups whose status is \emph{available}, as well as copying the values of $\transmit$-annotated variables to the new workgroups.

\paragraph{Accounting for concurrency}

Concurrent calls to $\offerfork$ must be protected by a mutex
to avoid racing on the
available workgroups. The same conflict over available workgroups can
occur at a new kernel launch between the scheduler and a workgroup
calling $\offerfork$.

The scheduler context provides an \emph{assignment mutes} to protect concurrent calls to
$\offerfork$ and $\offerkill$.  On calling $\offerkill$, a workgroup is only permitted to exit if the scheduler workgroup currently holds the assignment mutex; that is, the scheduler is trying to obtain work
groups to send to a task.  Conversely, on calling $\offerfork$ a workgroup can only spawn new workgroups if the scheduler workgroup does not hold the assignment mutex.  Thus the assignment mutex, only ever acquired by the scheduling workgroup, \emph{indirectly} provides mutual exclusion between a pairs of workgroups executing $\offerkill$ and $\offerfork$ concurrently. \ADComment{Please check previous -- I think I've got the coolness!}

Concurrent calls to $\offerkill$ are acceptable because only the
workgroup with the highest id may exit, automatically ensuring mutual
exclusion for exit-related modifications.

\paragraph{An efficient resizing barrier}

In \mysec\ref{sec:semantics}, we defined the semantics of a resizing barrier in terms
of a series of calls to $\globalbarrier$, $\offerfork$ and $\offerkill$.
 It is possible, however to implement the resizing with
only one call to a global barrier with $\offerfork$ and $\offerkill$
baked into the implementation. \TSComment{note here that these
  implementations require no cfork or ckill outside of the barrier} \ADComment{Ignore until camera-ready?}

We consider barriers that use the master/slave model
(e.g.\ \cite{XF10}): one workgroup (the master) collects signals
from the other workgroups (the slaves) indicating that they have
arrived at the barrier and are waiting for a reply indicating that
they may leave the barrier. Once the master has recieved a signal from
all the slaves, it replies with a signal saying that the slaves may
leave. \TSComment{Need to discuss new active workgroup number that
is stable across iterations} \ADComment{Again, ignore until camera-ready?}

A first-attempt at incorporating $\offerfork$ and $\offerkill$ into
such a barrier is straight forward. Upon entering the barrier, the
slaves first execute $\offerkill$, possibly exiting. The master then
waits for $M$ slaves (the number of active workgroups), which may
decrease due to $\offerkill$ calls by the
slaves, but will not increase. Once the master observes that $M$ slaves
have arrived, it knows that all other workgroups are spinning, waiting to be released. The master executes $\offerfork$,
and the statement immediately following this $\offerfork$ is a conditional that forces newly spawned workgroups to join the
slaves in waiting to be released. Finally, the master
releases all the slaves: the original slaves and the new slaves that
joined at $\offerfork$.

This barrier implementation is suboptimal becuase workgroups only
execute $\offerkill$ once per global barrier call and, depending on
the order in which workgroups arrive at the barrier, it is possible
that only one workgroup is killed per barrier call.  This may not
allow the scheduler to gather workgroups required to execute another
task in a timely fashion.

We can greatly reduce the gather time by providing a new $\keyword{query}()$
function for a workgroup executing a cooperative kernel, which returns the number of workgroups that the scheduler needs to obtained
from the cooperative kernel.  
%
The programmer cannot call $\keyword{query}$ directly;
it is only invoked from within the resizing barrier implementation.
%
A resizing barrier can now be implemented
as follows: (1) the master waits for all slaves to arrive; (2) the master calls $\offerfork$ and
commands the new workgroups to be slaves; (3) the master calls $\keyword{query}$, obtaining a value $W$;
(4) the master releases the slaves, broadcasting the value $W$ to them; (5) workgroups with ids larger than $M-W$ spin, calling 
$\offerkill$ repeatedly until the scheduler claims them---we know from $\keyword{query}$ that the scheduler will eventually do so.
We show in \mysec\ref{sec:responsiveness} that the barrier using
query greatly reduces the gather time in practice.

\section{Applications and Experiments}\label{sec:experiments}

We discuss our experience porting irregular algorithms to cooperative kernels (\mysec\ref{sec:portingalgorithms}) and discuss the GPUs on which we evaluate these applications (\mysec\ref{sec:gpus}).  For these GPUs, we report on experiments to determine non-cooperative workloads that model the requirements of various graphics rendering tasks (\mysec\ref{sec:sizingnoncoop}).  We then examine the overhead associated with moving to cooperative kernels when multitasking is \emph{not} required (\mysec\ref{sec:overhead}), as well as the responsiveness and throughput observed when a cooperative kernel is multi-tasked with non-cooperative workloads (\mysec\ref{sec:responsiveness}).

\subsection{Porting Irregular GPU Algorithms}\label{sec:portingalgorithms}

\begin{table}
\small
\centering
\begin{tabular}{ l r r r r r r}
App. & barriers & kill & fork & transmit & LoC & inputs\\
\hline
\rowcolor{Gray1}
color & 2 / 2 & 0 & 0 & 4 & 55 & 2\\
\rowcolor{Gray1}
mis & 3 / 3 & 0 & 0 & 0 & 71 & 2\\
\rowcolor{Gray1}
bc & 3 / 6 & 0 & 0 & 3 & 150 & 2\\
\rowcolor{Gray1}
p-sssp & 3 / 3 & 0 & 0 & 0  & 42 & 1\\
\rowcolor{Gray2}
bfs & 2 / 2 & 0 & 0  & 4  & 185 & 2\\
\rowcolor{Gray2}
l-sssp & 2 / 2 & 0 & 0  & 4  & 196 & 2\\
\rowcolor{Gray3}
octree & 0 / 0 & 1 & 1 & 0 & 213 & 1 \\
\rowcolor{Gray3}
game & 0 / 0 & 1 & 1 & 0 & 308 & 1 \\
\end{tabular} \\
\vspace{.2cm}
\crule[Gray1]{.2cm}{.2cm} Pannotia \hspace{.4cm} \crule[Gray2]{.2cm}{.2cm} Lonestar GPU  \hspace{.4cm}  \crule[Gray3]{.2cm}{.2cm} work stealing
\caption{Blocking GPU applications investigated}
\label{tab:applications}
\end{table}

\mytablong\ref{tab:applications} gives an overview of the 8 irregular
algorithms that we ported to cooperative kernels. Among them, 6 are
graph algorithms, based on Pannotia~\cite{Pannotia} and Lonestar~\cite{BNP12}
GPU application suites, using global barriers.  We indicate how many of
the original number of barriers are changed to resizing barriers, and
how many variables need to be transmitted.  In all cases all barriers
were converted, except for bc which contains barriers deeper in the call
stack, a case not supported by our prototype. Still, these barriers
could in principle also be converted. The remaining 2 algorithms are
workstealing applications: each required the addition of $\offerfork$
and $\offerkill$ at the start of the main loop, and no variables needed
to be transmitted.  For all these examples the porting to cooperative
kernels was similarly straightforward to the porting illustrated in
\mysec\ref{sec:programmingguidelines} for our running examples, and did
not require reshaping of the original algorithm. Most graph come with 2
different data sets, leading to 13 application/data set pairs in total.

From the Lonestar suite we omit two applications that use global barriers:
dmr has been shown to be unstable across GPU vendors~\cite{DBLP:conf/iwocl/SorensenD16}, and mst is trivially short-running so that it is not
interesting for our multitasking approach.
%
These exceptions aside, we have handled all the available OpenCL
implemenatinos of irregular algorithms that we are aware of, plus two
easy-to-adapt CUDA algorithms (the workstealing examples).

\subsection{The GPUs we Tested}\label{sec:gpus}

Our prototype implementation (\mysec\ref{sec:implementation}) requires
two optional features of OpenCL 2.0: SVM fine-grained buffers and SVM
atomics. This rules out evaluation on \nvidia and ARM Mali GPUs at
present, since these vendors do not yet support OpenCL $2.0$,
equivalent functionality is not provided by \nvidia's CUDA. Intel
drivers meet our requirements on Windows, but not on Linux (we
confirmed this by trying the upstream driver that requires a patched
Linux kernel). AMD support SVM atomics only their Kaveri family of
chips and only on Linux, requiring to compile the latest driver.

Among the GPUs available to us, four met our
requirements: Intel HD 520, HD 5500 and Iris 6100, and AMD Radeon R7.
However, AMD's Linux drivers for OpenCL suffer from a known defect
whereby long-running kernels lead to defunct processes that the OS
cannot kill.\footnote{See \url{http://developer.amd.com/download/AMD_APP_SDK_Release_Notes_Developer.pdf}}  We observed this on our AMD platform with the latest
driver, and simultaneously obtained alarming results on this platform
when using SVM atomics: modifying a kernel to include a simple
host/device handshake through SVM slowed down kernel execution by an
order of magnitude in some cases.  This behaviour, combined with the
defunt processes issue, lead us to suspect that the support we require
for cooperative kernels is not yet mature enough within AMD's drivers,
which we plan to discuss with engineers at AMD.

We thus present experimental results for the Intel GPUs, whose characteristics are summarised in \mytab\ref{tab:chipstested}.

\begin{table}
\footnotesize
\centering
\begin{tabular}{ l l l l }
Chip & HD520 & HD5500 & Iris\\
\hline
\# CU  & 24 & 24 & 47 \\
Driver   & 20.19.15.4501 & 10.18.15.4281 & 20.19.15.4463 \\
Host CPU & i3-6100U & i7-5600U & i3-5157U \\
\end{tabular}
\caption{The Intel GPU platforms used in our evaluation}
\label{tab:chipstested}
\end{table}

\subsection{Sizing Non-Cooperative Kernels}\label{sec:sizingnoncoop}

\ADComment{Refer back to nonfunctional requirements section, \mysec\ref{sec:nonfunctional}.}

Enabling rendering of smooth graphics in parallel with execution of irregular algorithms is an important use case for our approach.  Because our prototype implementation is based on a mega-kernel that takes over the entire GPU (see \mysec\ref{sec:implementation}), we cannot assess this use case directly.

As a proxy for genuine graphics rendering, we devised the following
procedure to determine non-cooperative OpenCL workloads that simulate
the computational intensity of various graphics rendering workloads.
%
We designed a synthetic kernel that occupies all workgrouops of a
given GPU for a parameterised length of time $t$.  We wrote a host
program to repeatedly launch this kernel in an infinite loop.  We then
searched for a maximum value for the parameter $t$ that allowed the
synthetic kernel to execute without having an observable impact on
graphics rendering, where ``observable'' was decided manually by one
of the paper authors.  Having found the appropriate value for $t$, we
ran the host application for $X$ seconds, measuring the time $Y < X$
dedicated to GPU execution during this period, as well as the number
of kernel launches $n$ that were issued.  We used $X \geq 10$ in all
experiments.  The values $(X-Y)/n$ and $X/n$ estimate the average time
spent using the GPU to render the display between kernel calls and the
rate at which the OS requires the GPU for display rendering,
respectively.

We used this approach to measure the GPU availability required for
three types of rendering: \emph{light}, whereby desktop icons are
smoothly emphasised under the mouse pointer; \emph{medium}, whereby
window dragging over the desktop is smoothly animated; and
\emph{heavy}, which requires smooth animation of a WebGL shader in a
browser.  For \emph{heavy} we used the Paper
Planes\footnote{\url{https://paperplanes.world}}
and Kaleidoscope\footnote{\url{http://colordodge.com/Kaleidoscope/}}
shaders running in a recent version of Firefox.

\begin{table}
\small
\centering
\begin{tabular}{ l r r }
Workload & Duration (ms) & Rate (ms)\\
\hline
light & 3 & 70\\
medium & 3 & 40\\
heavy & 10 & 40\\
\end{tabular}
\caption{Computed estimates of the GPU requirements for three different classes of graphics rendering}
\label{tab:noncooperativeworkload}
\end{table}

\mytablong\ref{tab:noncooperativeworkload} illustrates the data
computed for the three types of workloads across our GPUs.  The
\emph{Duration} and \emph{Rate} columns report values corresponding to
$(X-Y)/n$ and $X/n$ in the above discussion, respectively.  For the
medium and heavy workloads, the 40ms rate corresponds to graphics
rendering occuring 25 times per second, which coincides with the human
persistence of vision. The 3ms execution duration of both light and
medium configurations indicates that, as expected, GPU computation is
cheaper for basic display rendering compared with rendering that uses
complex shaders.

We use these results to guide our evaluation of the responsiveness of
cooperative scheduling in \mysec\ref{sec:responsiveness}.

\subsection{The Overhead of Moving to Cooperative Kernels}\label{sec:overhead}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% chip: hd520
% overall - 
% geomean: 1.19835906125
% max ('color', 'ecology', 1.749620120141443)
% ----- 
% barrier 
% max - ('color', 'ecology', 1.749620120141443)
% geomean: 1.18027192048
% ----- 
% workstealing
% max - ('octree', 'NA', 1.4165788173996423)
% geomean: 1.4165788174
% ----- 
% chip: iris
% overall - 
% geomean: 1.07926294253
% max ('sssp', 'usa_ny', 1.3661910830147503)
% ----- 
% barrier 
% max - ('sssp', 'usa_ny', 1.3661910830147503)
% geomean: 1.07122874447
% ----- 
% workstealing
% max - ('octree', 'NA', 1.2333896396396398)
% geomean: 1.11616906144
% ----- 
% chip: hd5500
% overall - 
% geomean: 1.15074583378
% max ('bc', '1k_128k', 1.4466039296624815)
% ----- 
% barrier 
% max - ('bc', '1k_128k', 1.4466039296624815)
% geomean: 1.16116637952
% ----- 
% workstealing
% max - ('octree', 'NA', 1.1814511945151949)
% geomean: 1.1049985913
% ----- 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
\small
\centering
\begin{tabular}{ l l l l }
Chip & HD520 & HD5500 & Iris\\
\hline
overall geo. mean  & $1.20$ & $1.15$ & $1.08$ \\
overall max & $1.75^{\ast}$ & $1.45^{\dagger}$ & $1.37^{\ddagger}$ \\

\hline
barrier geo. mean & $1.18$ & $1.16$ & $1.07$ \\
barrier max & $1.75^{\ast}$ & $1.45^{\dagger}$ & $1.37^{\ddagger}$ \\

\hline
wk.steal. geo. mean & $1.42$ & $1.10$ & $1.12$ \\
wk.steal. max & $1.42^{\diamond}$ & $1.18^{\diamond}$ & $1.23^{\diamond}$ \\

\hline
\end{tabular}\\
{\footnotesize
$^{\ast}$color ecology, $^{\dagger}$bc 1k\_128k, $^{\ddagger}$sssp usa\_ny, $^{\diamond}$octree
}
\caption{Execution time slowdown associated with cooperative scheduling mode without multitasking}
\label{tab:overhead}
\end{table}

\paragraph{Experimental setup}
Invoking the cooperative scheduling primitives incurs some overhead
even if no killing, forking or resizing actually occurs, because the cooperative kernel still needs to interact with the scheduler to determine this.
We assess this overhead by measuring the
 slowdown in execution time between the original and cooperative versions of a kernel, forcing the scheduler to never modify the number of
active workgroups in the cooperative case.

Recall that our mega kernel-based implementation merges the code for a
cooperative kernel with that of the non-cooperative task to be
co-scheduled.  This blow-up in code size can have a negative impact on
compilation, e.g.\ due to higher register pressure, reducing the
occupancy for the merged kernel---the maximum number of workgroups
that the GPU driver will schedule.  This is an artifact of our
proof-of-concept implementation strategy, and would not be a problem
if our cooperative scheduling approach was implemented inside the GPU
driver.  We thus launch both the original and cooperative versions of
a kernel with the reduced occupancy bound in order to meaningfully
compare execution times.

\paragraph{Results}
For each GPU, \mytab\ref{tab:overhead} shows the geometric mean and
maximum slowdown across all applications and data sets, with averages and maxima computed over 10 runs per benchmark. For the maximum slowdowns, we indicate which
application and data set was responsible.  The slowdown is below 2 even
in the worst case, and closer to 1 on average. We consider these results
as encouraging, especially since the performance of our prototype could
clearly be improved upon if cooperative scheduling was implemented
directly in the GPU driver.

\ADComment{Can we remark a bit more on the outliers?}

\subsection{Multitasking via Cooperative Scheduling}\label{sec:responsiveness}

We now assess the performance impact of multitasking between a
coopoerative and non-cooperative kernel, and on scheduling
responsiveness.  The performance is measured with respect to the
performance of each kernel run in sequence. Responsiveness, i.e., how
fast a non-cooperative kernel can be scheduled, depends not only on
the scheduler implementation, but also on how often the cooperative
kernel offer workgroups to be killed.

% In order to simulate non-persistent workloads alongside kernels of
% blocking algorithms, we use a matrix multiplication kernel where
% matrix dimensions are tuned to trigger the relevant execution
% duration.

\paragraph{Experimental setup}
For a given cooperative kernel and its data set, we launch the kernel
asynchronously and then repeatedly schedule a non-cooperative kernel
that aims to simulate the intensity of a graphics rendering workload,
according to the classes listed in
\mytab~\ref{tab:noncooperativeworkload}. In practice, we use a matrix
multiplication kernel as the non-cooperative workload, with matrix
dimensions tailored for each GPU to reach the appropriate execution
duration.  We conduct separate runs where we vary the number of workgroup slots requested by the non-cooperative kernel,
considering the cases where one, a quarter, a half, and all-but-one, of the total number of workgroup slots is requested.
Moreover, for the graph algorithms we try both the
regular and query barrier implementations.

Our experiments span 13 pairs of cooperative kernels and data sets, 3
classes of non-cooperative kernel workloads \ADComment{cut ``and frequency''?} and frequency, 4 quantities
of workgroups claimed for the non-cooperative kernel and 2 variations of
resizing barriers for graph algorithms, leading to 288
configurations.  We run each configuration 10 times on each
GPU, in order to report averaged performance numbers. For each run, we record the execution time of the cooperative
kernel. For each scheduling of the non-cooperative kernel during the
run, we also record the \emph{gather time} needed by the scheduler to
collect enough workgroups to launch the non-cooperative kernel, and the
non-cooperative kernel execution time. We sanity-check whether all kernels perform
their computations corretly by comparing their computed result with expected reference results.

To work around compiler crash and compiler hang bugs in Intel's current OpenCL drivers, we had to apply what semantics-preserving changes to the code generated by \kernelmerge{} for a few applications configurations.  We could not obtain results for some configurations where the non-cooperative kernel asks
for all but one workgroups (namely bc, bfs and l-sssp on all chips, and
color on Iris), and could not get the game application to run at all on HD520; these configurations led to 
catastrophic issues such as display freezes or blue screens; we also attribute these to bugs in Intel's current drivers.

% Whenever a cooperative or non-cooperative kernel terminates, we verify
% that it has really performed its computation by comparing its results
% against the expected one. We encountered two kind of issues during the
% experiments: some configurations would lead to a system freeze or crash
% although they run fine on other GPUs. Although we don't claim that our
% programs are all correct, this issue seems to indicate a bug in the
% OpenCL compiler or driver of the problematic GPU. We also have some
% cooperative kernels that returned invalid results when the
% non-cooperative kernel takes all but one of the cooperative kernel
% workgroups, we decided to ignore these results.

\paragraph{Results}

\myfiglong\ref{fig:fine-grained-timing} reports the measure for three
configurations and represents the gather and execution time of the
non-cooperative kernel w.r.t. the quantity of workgroups allocated to
it. Times are in milliseconds, and a logarithmic scale is used since
gather times tend to be much smaller than execution times. The
horizontal grey lines indicates the desired rate for non-cooperative
kernels.

Responsiveness of the cooperative scheduling is measured by the gather
time. In all examples, we observe that the gather time grows as we ask
for more workgroups for the non-cooperative kernel: this is expected,
since the scheduler must wait for more workgroups to call $\offerkill$
or to reach a resizing barrier. Still, notice that the query barrier
leads to better and stable gather times for graph applications, since it
can collect all needed workgroups at the first time they pass through
the barrier. Gather time also application-dependent since it is related
to how often the cooperative kernel workgroups hit $\offerkill$ or
$\resizingglobalbarrier$.

\HEComment{missing results for all but one in bfs, discussed in expe
setup, do we need to say it here ?}

\begin{figure*}
\includegraphics[width=.7\columnwidth]{iris_octree_NA.pdf} 
\includegraphics[width=.7\columnwidth]{hd520_mis_ecology.pdf}
\includegraphics[width=.7\columnwidth]{hd5500_bfs_rmat22.pdf}
\caption{Example gather time and non-cooperative timing results}\label{fig:fine-grained-timing}
\end{figure*}

\HEComment{Figure gather time: replace ``O'' by ``M'' , ie max nb of active workgroups}

\HEComment{Figure gather time: add ``ms'' to time coordinate}

\HEComment{Figure gather time: lower legend in hd5500 bfs figure}

\HEComment{Figure gather time: use consistent line styles for work stealing (ie, same style for ``heavy'' across all figures)}

\section{Related Work}\label{sec:relatedwork}

\paragraph{Irregular algorithms and persistent kernels}

There has been a lot of work on accelerating blocking irregular
algorithms using GPUs, and on the \emph{persistent threads}
programming style for long-running
kernels~\cite{owens-persistent,DBLP:conf/ipps/KaleemVPHP16,DBLP:conf/ipps/DavidsonBGO14,DBLP:conf/hipc/HarishN07,DBLP:journals/topc/MerrillGG15,DBLP:conf/egh/VineetHPN09,DBLP:conf/ppopp/NobariCKB12,DBLP:conf/hpcc/SolomonTT10a,DBLP:conf/popl/PrabhuRMH11,DBLP:conf/ppopp/Mendez-LojoBP12,DBLP:conf/oopsla/PaiP16,DBLP:conf/oopsla/SorensenDBGR16,DBLP:conf/egh/CedermanT08,TPO10,BNP12,Pannotia}.
These approaches rely on the occupancy-bound execution model, flooding
available compute units with work, so that the GPU is unavailable for
other tasks, and assuming fair scheduling between occupant workgroups,
which is unlikely to be guaranteed on future GPU platforms.
%
As our experiments demonstrate, our cooperative kernels model allows blocking algorithms
to be upgraded to run under a fair scheduler in a manner that, with
appropriate care, facilitates responsive multitasking.

\paragraph{Preemptive multitasking on GPUs}

Hardware support for preemption has been proposed for \nvidia GPUs,
as well as \emph{SM-draining}
whereby workgroups occupying a symmetric multiprocessor (SM; a compute
unit using our terminology) are allowed to complete until the SM becomes free for
other tasks~\cite{DBLP:conf/isca/TanasicGCRNV14}.  SM draining is limited the presence of blocking constructs, since it may not be possible to drain a blocked workgroup.
%
A follow-up work adds the notion of SM \emph{flushing},
where a workgroup can be cancelled and re-scheduled
from scratch if it has not yet committed side-effects~\cite{DBLP:conf/asplos/ParkPM15}.  Both approaches have been evaluated
using simulators, over sets of regular GPU kernels.  Very recent \nvidia GPUs targeting data centers now support
preemption, though it is not clear whether they guarantee fair scheduling~\cite{PascalWhitepaper}.

The advantages of preemption are that it does not require programmer
effort, and guards against a rogue long-running kernel making the GPU
unavailable for other tasks.  The main drawback is that \emph{every}
kernel takes a potential performance hit.
%
If implemented natively, our solution does not impact performance on regular kernels that are not scheduled with cooperative kernels.
Furthermore, careful placement of
our new primitives can exploit problem structure, enabling more efficient multitasking than
application-oblivious preemption.  There is scope for combining the approaches: if a
cooperative kernel is insufficiently cooperative, i.e.\ it does not
invoke $\offerkill$ enough, then hardware support for preemption could
be invoked to prevent the kernel from denying other tasks access to
the GPU indefinitely.

\paragraph{Optimizing GPU scheduling}
%
A number of works have considered how best to schedule dynamic
workloads on GPUs~\cite{DBLP:journals/tog/SteinbergerKKHKS12,DBLP:conf/ics/WuCLSV15,DBLP:conf/usenix/KatoLRI11,DBLP:journals/tog/SteinbergerKBKDS14,DBLP:journals/rts/ElliottA12,DBLP:conf/sosp/RossbachCSRW11}.  Among these, the Whippletree approach employs a persistent
megakernel to schedule multiple, dynamic tasks in a manner that aims
to best utilise GPU
resources~\cite{DBLP:journals/tog/SteinbergerKBKDS14}, and the TimeGraph
approach similarly aims to optimise scheduling of competing workloads,
in the context of OpenGL~\cite{DBLP:conf/usenix/KatoLRI11}.
None of these approaches tackle
the problem of \emph{fair} scheduling on GPUs, thus they do not aid in safe deployment of blocking irregular algorithms.

CUDA and OpenCL provide the facility for a kernel to spawn further kernels~\cite{cuda-75,opencl2Spec}.
This \emph{dynamic parallelism} can be used to implement a
GPU-based scheduler, by having an initial scheduler kernel repeatedly
spawn further kernels as required, according to some scheduling
policy~\cite{DBLP:conf/ppopp/Muyan-OzcelikO16}.  However, kernels
that uses dynamic parallelism are still prone to unfair scheduling of workgroups, so this feature
does not help solve the problem of safely deploying blocking
algorithms on GPUs.

\paragraph{Cooperative multitasking}

Cooperative multitasking was offered in older operating systems
(e.g. Windows before the 95 edition) and is still used by some
operating systems, such as RISC
OS.\footnote{\url{http://www.riscos.info/index.php/Preemptive_multitasking}}
Many programming languages offer cooperative multitasking via
variations of coroutines, e.g. coroutine objects in
Python
%\footnote{\url{https://docs.python.org/3/reference/datamodel.html\#coroutine-objects}}
and Lua,
%\footnote{\url{https://www.lua.org/pil/9.1.html}}
and Ruby's
\emph{fiber}.
%\footnote{\url{http://ruby-doc.org/core-2.1.1/Fiber.html}}

In the context of GPUs, cooperative multitasking has been used to
refer to the standard programming model, where the onus is on the GPU kernel
to complete execution within a reasonable time budget~\cite{adriaens2012case,CPE:CPE1722}.
%
This is in contrast to our cooperative kernels, which specifically aim to support the needs of long-running GPU tasks.

\section{Conclusions and Future Work}\label{sec:conclusion}

We have proposed \emph{cooperative kernels}, a small set of GPU
programming extensions that allow long-running, blocking kernels to be
fairly scheduled and to share GPU resources with other workloads.
Experimental results using our megakernel-based prototype over three
Intel GPUs shows that the model is a good fit for the needs of current
GPU-accelerated irregular algorithms.  The
performance that could be gained through a native implementation with
dedicated driver support would be even better.

There are three main avenues for future work at present: seeking
additional classes of irregular algorithms to which the model might
(be extended to) apply (to), investigating native driver support for
cooperative kernels via open source graphics drivers, and integrating
cooperative kernels into template- and compiler-based programming
models for graph algorithms on
GPUs~\cite{SreesCompetition,DBLP:conf/oopsla/PaiP16}.  As more vendors
adopt the advanced features of OpenCL that our approach requires, we
will be in a position to assess the programming model across a wider
range of devices.

\clearpage

\bibliographystyle{abbrvnat}
\bibliography{tyler}

\clearpage

\appendix

\section{Operational Semantics for Cooperative Kernels}\label{appendix:semantics}

\newcommand{\myss}{\mathit{ss}}
\newcommand{\Stmts}{\mathsf{Stmts}}
\newcommand{\threadstates}{\mathsf{ThreadStates}}
\newcommand{\sharedstates}{\mathsf{SharedStates}}
\newcommand{\sync}{\mathsf{sync}}

In \mysec\ref{sec:semantics} we presented the semantics of cooperative
kernels relatively informally, using English, to provide the intuition
behind our programmming model.  We now back this up with a more formal
presentation as an operational semantics for an abstract GPU
programming model.

\paragraph{States}

Let $L$ be a set of \emph{local states} that abstractly captures the
private memory associated with a thread executing a GPU kernel.  Let
$\Stmts$ denote the set of all possible statements that a
thread can execute.  We do not detail the structure of these
statements, except that we assume sequential composition of statements
is provided by the $\code{;}$ separator, and that the $\offerkill$,
$\offerfork$, $\globalbarrier$ and $\resizingglobalbarrier$ primitives
from our cooperative kernels programming model are valid statements.

A \emph{thread state} is then a pair $(l, \myss)$, where $l \in L$ and
$\myss \in \Stmts$.  The $l$ component captures the valuation of all the
thread's private memory, and the $\myss$ component captures the
remaining statements to be executed by the thread.  Let $\threadstates$ denote the set of all thread states.

Assuming that $d > 0$ threads per workgroup were requested on kernel launch, a \emph{workgroup state}
is a $d$-tuple $((l_0, \myss_0), \dots, (l_{d-1}, \myss_{d-1}))$, where each $(l_i, \myss_i)$ is the thread state for the $i$th thread in the workgroup $(0\leq i < d$).

Assuming that $N > 0$ was specified as the maximum number of
workgroups that should execute the cooperative kernel, a \emph{kernel
  state} is then a pair
%
\[
(\sigma, (w_0, \dots, w_{M-1}, \underbrace{\bot, \dots,
\bot}_{N-M}))\]
%
where: $\sigma$ represents the shared state of the kernel; $M \leq N$
is the number of \emph{active} workgroups; $w_i$ is the workgroup
state for active workgroup $i$ ($0 \leq i < M$); and $N-M$ occurrences
of $\bot$ indicate \emph{absent} workgroups.  Let $\sharedstates$
denote the set of all possible shared states.  We regard
workgroup-local storage as being part of the shared state of a kernel.

\paragraph{Thread-level transitions}

We leave the semantics for thread-level transitions abstract, assuming
a binary relation $\rightarrow_{tau}$ on $\sharedstates{} \times
\threadstates{}$.  If $(\sigma, (l, \myss)) \rightarrow_{\tau}
(\sigma', (l', \myss'))$, this indicates that if a thread is in local
state $(l, \myss)$, the thread can transition to local state $(l',
\myss')$, changing the shared state from $\sigma$ to $\sigma'$ in the
process.

All we require is that $(\sigma, (l, \myss)) \rightarrow_{\tau}
(\sigma', (l', \myss'))$ if $\myss$ has the form $\mathsf{special}();
\mathit{tt}$, where $\mathsf{special}$ is one of $\offerkill$,
$\offerfork$, $\globalbarrier$ or $\resizingglobalbarrier$.  This is
because we shall specifically define the meaning of the new primitives
introduced by our programming model.

\paragraph{Memory synchronisation}

GPUs are known to have relaxed memory models~\cite{ABDGKPSW-2015}.  To abstractly
account for this, we assume that the shared state component $\sigma$
is not simplyl a mapping from locations to values, but instead
captures all the intricacies of GPU memory that can lead to this
relaxed behaviour.  We also assume a function $\sync$ which, given a
kernel state $\kappa$, returns a set of kernel states.  The idea is
that each $\kappa' \in \sync(\kappa)$ is a possible kernel state that
can be reached from $\kappa$ by stalling until all stores to memory
and loads from memory to thread-local storage have completed.  All we
require is that $\sync$ does not modify the number of active
workgroups nor the component of a thread state that determines which
statements remain to be executed.

\begin{figure*}
\begin{center}

\[
\inferrule{
w_i(j) = (l, \myss)
\\
(\sigma, (l, \myss)) \rightarrow_{\tau} (\sigma', (l', \myss'))
\\
w_i' = w_i[j \mapsto (l', \myss')]
}
{
(\sigma, (\dots, w_i, \dots)) \rightarrow (\sigma', (\dots, w_i', \dots))
}
(\textsc{Thread-Step})
\]

\medskip

\[
\inferrule{
\forall j \;.\; w_i(j) = (l_j, \offerkill();\myss)
\\
w_i' = ((l_0, \myss), \dots, (l_{d-1}, \myss))
}
{
(\sigma, (\dots, w_i, \dots)) \rightarrow (\sigma, (\dots, w_i', \dots))
}
(\textsc{Kill-No-Op})
\]

\medskip

\[
\inferrule{
\forall j \;.\; w_{M-1}(j) = (l_j, \offerkill();\myss)
\\
M > 0
}
{
(\sigma, (\dots, w_{M-2}, w_{M-1}, \bot, \dots, \bot)) \rightarrow (\sigma, (\dots, w_{M-2}, \bot, \bot, \dots, \bot))
}
(\textsc{Kill})
\]

\medskip

\[
\inferrule{
\forall j \;.\; w_i(j) = (l_j, \offerfork();\myss)
\\
w_i' = ((l_{0}, \myss), \dots, (l_{d-1}, \myss))
\\
k \in [0, N - M]
\\
\forall a \in [0, k - 1] \;.\; w_{M+a} = ((l_{0}, \myss), \dots, (l_{0}, \myss))
}
{
(\sigma, (\dots, w_i, \dots, w_{M-1}, \bot, \dots, \bot)) \rightarrow (\sigma, (\dots, w_i', \dots, w_{M-1}, w_{M}, \dots, w_{M+k-1}, \bot, \dots, \bot))
}
(\textsc{Fork})
\]

\medskip

\[
\inferrule{
\forall i \;.\;\forall j\;.\;w_i(j) = (l_{i,j}, \globalbarrier();\myss)
\\
\forall i \;.\;\forall j\;.\;w_i'(j) = (l_{i,j}, \myss)
\\\\
\kappa \in \sync((\sigma , (w_{0}', \dots, w_{M-1}', \bot, \dots, \bot))
}
{
(\sigma, (w_{0}, \dots, w_{M-1}, \bot, \dots, \bot)) \rightarrow \kappa
}
(\textsc{Barrier})
\]

\medskip

\[
\inferrule{
\forall i \;.\;\forall j\;.\;w_i(j) = (l_{i,j}, \resizingglobalbarrier();\myss)
\\
\forall j\;.\;w_{0}'(j) = (l_{1,j}, \globalbarrier(); \offerfork(); \globalbarrier(); \globalbarrier();\myss)
\\
\forall i \neq 1 \;.\;\forall j\;.\;w_i'(j) = (l_{i,j}, \globalbarrier(); \globalbarrier(); \offerkill(); \globalbarrier();\myss)
}
{
(\sigma, (w_{0}, \dots, w_{M-1}, \bot, \dots, \bot)) \rightarrow (\sigma, (w_{0}', \dots, w_{M-1}', \bot, \dots, \bot))
}
(\textsc{Resizing-Barrier})
\]

\end{center}

\caption{Abstract operational semantics for our cooperative kernels language extensions}\label{fig:semanticrules}

\end{figure*}

\paragraph{Operational semantics rules}

\myfiglong\ref{fig:semanticrules} presents the rules of our
operational semantics, defining a relation $\rightarrow$ on kernel
states.

Rule $\textsc{Thread-Step}$ defines the semantics for thread making a
single execution step, delegating to the abstract $\rightarrow_{\tau}$
relation to determine how the thread's local state and the shared
state component change.  For simplicity, this rule ignores the
semantics of intra-workgroup barriers, which are not our focus here.

Rule $\textsc{Kill-No-Op}$ reflects the fact that when a workgroup
reaches $\offerkill$, this can always be treated as a no-op.  Whether
a scheduler implementation accepts $\offerkill$ calls or not depends
on competing workloads and how the scheduler has been designed to meet
the non-functinoal requirements discussed in
\mysec\ref{sec:nonfunctional}, but in general the programmer should
always be prepared for the possiblity that a workgroup survives after
calling $\offerkill$.

The case where a workgroup's offer to be killed is accepted by the
scheduler is captured by rule $\textsc{Kill}$.  Because we have
adopted a semantics where workgroup 0 is never killed and where only
the workgroup with the highest id can be killed, the rule only fires
if $M > 0$ and workgroup $M-1$ has reached $\offerkill$.  The rule has
the effect of replacing the workgroup state for $w_{M-1}$
with $\bot$.

Recall that $\offerkill$ is a workgroup-level function: the same
syntactic $\offerfork$ call must be reached by all threads in a
workgroup.  This is captured in our rules by requiring in both
$\textsc{Kill-No-Op}$ and $\textsc{Kill}$ that every thread is ready
to execute $\offerkill$ followed by an identical statement $\myss$.
Neither rule applies until all threads in a workgroup reach
$\offerfork$, and the workgroup gets stuck if multiple threads in a
workgroiup reach $\offerfork$ with different subsequent statements.

The $\textsc{Fork}$ rule similarly requires all threads in a workgroup
to reach $\offerfork$ with identical following statements.  A
nondeterministic number of new workgroups, $k$, is selected to be
forked, where $k \in [0, N-M]$.  Importantly, $k=0$ is always a
legitimate choice, in which case $\offerfork$ has no effect on the
number of workgroups that are executing.  In the case where $k > 0$,
$k$ new workgroup states are created, where each workgroup inherits
the local state of thread 0 in $w_i$, the workgroup executing the fork
call.  After $\offerfork$, all threads in all workgroups, including
the new threads, proceed to execute the sequence of statements $\myss$
that followed $\offerfork$.

A simplification here is that we do not model transmission of
particular annotated variables from thread 0 of the forking workgroup,
instead specifying that the entire local state of the workgroup is
transmitted.  Extending the semantics to transmit only annotated
variables would be straightforward but verbose, requiring the local
memory component of a thread state to be split into two: the part of
the local state to be transmitted (modelling $\transmit$-annotated
variables), and the rest of the local state.

The $\textsc{Barrier}$ rule requires that \emph{every} thread
executing the kernel reaches $\globalbarrier()$ an identical following
statement.  This reflects that fact that $\globalbarrier()$ is a
kernel-level function.  Each thread then skips over the
$\globalbarrier()$ call, and the $\sync$ function is applied to yield
a set of kernel states that can arise due to memory synchronization
taking place.  An arbitary member of this set is selected as the next
kernel state.

Despite its apparent complexity, the $\textsc{Resizing-Barrier}$ rule simply implements the rewriting of $\resizingglobalbarrier$ in terms of $\globalbarrier$, $\offerkill$ and $\offerfork$ discussed in \mysec\ref{sec:semantics}.

\section{Alternative Semantic Choices}\label{appendix:semanticalternatives}

The semantics of cooperative kernels has been guided by the practical
applications we have studied (described in
\mysec\ref{sec:portingalgorithms}).  We now discuss several cases
where we might have taken different and also reasonable semantic
decisions.

\paragraph{Killing order}

We opted for a semantics whereby only the active workgroup with the
highest id can be killed.  This has an appealing property: it means
that the ids of active workgroups are contiguous, which is important
for processing of contiguous data.  The cooperative graph traversal
algorithm of \myfig\ref{fig:cgraphsearch} illustrates this: the
algorithm is prepared for $\getglobalsize()$ to change after each
resizing barrier call, but depends on the fact that $\getglobalid()$
returns a contiguous range of thread ids.

A disadvantage of this decision is that it may provide sub-optimal
responsiveness from the point of view of the scheduler.  Suppose the
scheduler requires an additional compute unit, but the active thread
with the larget id is processing some computationally intensive work
and will take a while to reach $\offerkill$.  Our semantics means that
the scheuduler cannot take advantage of the fact that another active
workgroup may invoke $\offerkill$ sooner.

Cooperative kernels that do not require contiguous thread ids might me
more suited to a semantics in which workgroups can be killed in any
order, but where workgroup ids (and thus thread global ids) are not
guaranteed to be contiguous.

\paragraph{Keeping one workgroup alive}

Our semantics dictate that the workgroup with id 0 will not be killed
if it invokes $\offerkill$.  This avoids the possibility of the
cooperative kernel terminating early due to the programmer
inadvertantly allowing all workgroups to be killed, and the decision
to keep workgroup 0 alive fits well with our choice to kill workgroups
in descending order of id.

However, there might be a use case for a cooperative kernel reaching a
point where it would be acceptable for the kernel to exit, although
desirable for some remaining computation to be performed if competing
workloads allow it.  In this case, a semantics where all workgroups can be killed via $\offerkill$ would be appropriate, and the programmer would need to guard each $\offerkill$ with an id check in cases where killing all workgroups would be unacceptable.  For example:
%
\lstset{basicstyle=\tt,numbers=none}
\begin{lstlisting}
  if($\getgroupid{0}$ != 0) $\offerkill$();
\end{lstlisting}
\lstset{basicstyle=\scriptsize\tt,numbers=left}
%
would ensure that at least workgroup 0 is kept alive.

\paragraph{Transmission of partial state from a single thread}

Recall from the semantics of $\offerfork$ that newly forked workgroups
inherit the variable valuation associated with thread 0 of the forking
workgroup, but only for $\transmit$-annotated variables.  Alternative
choices here would be to have forked workgroups inherit values for
\emph{all} variables from the forking workgroup, and to have thread
$i$ in the forking workgroup provide the valuation for thread $i$ in
each spawned workgroup, rather than having thread 0 transmit the
valuation to all new threads.

We opted for transmitting only selected varibles based on the
observation that many of a thread's private variables are dead at the
point of issuing $\offerfork$ or $\resizingglobalbarrier$, thus it
would be wasteful to transmit them.  A live variable analysis could
instead be employed to over-approximate the variables that might be
accessed by newly arriving workgroups, so that these are automatically
transmitted.

In all cases, we found that a variable that needed to be transmitted
had the property of being uniform across the workgroup.  That is,
despite each thread having its own copy of the variable, each thread
is in agreement on the variable's value.  As an example, the
$\keyword{level}$, $\keyword{in\_nodes}$ and $\keyword{out\_nodes}$
variables used in \myfig\ref{fig:cgraphsearch} are all stored in thread-private
memory, but all threads in a workgroup agree on the values of these
variables at each $\resizingglobalbarrier$ call.  As a result,
transmitting the thread 0's valuation of the annotated variables is
equivalent to (and more efficient than) transmitting values on a
thread-by-thread basis.  We have not yet encountered a real-world
example where our current semantics would not suffice.

\section{Full Set of Graphs for Responsiveness Experiments}

\ADComment{TODO: we need this.  Probably not the right title.}

\end{document}